---
title: "COVID-19 in Switzerland and Western Europe"
output:
  html_notebook: 
    fig_height: 5
    fig_width: 7
    number_sections: yes
    toc: yes
  html_document: 
    df_print: default
    number_sections: yes
    toc: yes
  word_document: 
    toc: yes
  pdf_document: default
date: "Last updated: `r format(Sys.time(), '%d %B, %Y - %H:%M')`"
author: "Philippe Docourt"
---

```{r, echo=FALSE, warning=FALSE}
library(data.table)

get_timeseries_as_vector <- function(df) {
  tmp <- subset(df, select = -c(Province.State, Lat, Long))
  tmp <- sapply(as.vector(transpose(tmp)), as.integer)
  return(tmp)
}

plot_time_series_for_country <- function(country.confirmed, country.deaths, country.recovered, country.name, public.spaces.closed.at=NA, confinement.at=NA) {
    day.count <- nrow(country.confirmed)
  ylim <- c(0, 1.3 * max(country.confirmed, na.rm = TRUE))
  day.start <- min(c(35, public.spaces.closed.at, confinement.at), na.rm = TRUE)
  plot(country.confirmed, main= paste("Cases in", country.name), type = "o", xlab = "Days Since January 21, 2020", ylab = "Number of Cases", col = "red", xlim =  c(day.start, day.count + 7), ylim = ylim)
  lines(seq(1, length(country.deaths)), country.deaths, type = "l", col = "black", lty = 1, lw = 1)
  lines(seq(1, length(country.recovered)), country.recovered, type = "l", col = "green", lty = 1, lw = 1)
  grid()
  axis(side = 1, at = seq(day.start %/% 5 * 5, day.count, 5))
  
  # Trend based on last two observations (last day).
  last.confirmed <- tail(country.confirmed, n = 2)
  dy <-last.confirmed[2] - last.confirmed[1]
  slope <- dy
  u <- c(1, dy)
  a <- last.confirmed[1] - (day.count - 1) * slope
  abline(a = a, b = slope, col = "violet", lw = 1.5, lty = 4)
  last.progression.pct <- round((100 * dy) / last.confirmed[1])
  text(x = day.count - 4, y = 0.5 * ylim[2], paste(dy, "cases/day\nExtrapolation:", dy * 7, "case/week\n", last.progression.pct, "% progress last day"), srt = 20, col = "violet")
  
  # Trend based on last 8 observations (7 intervals, one week)
  last.confirmed <- tail(country.confirmed, n = 8)
  dy <-last.confirmed[8] - last.confirmed[1]
  slope <- dy / 7
  u <- c(7, dy)
  a <- last.confirmed[1] - (day.count - 7) * slope
  abline(a = a, b = slope, col = "black", lw = 1.5, lty = 4)
  last.progression.pct <- round((100 * dy) / last.confirmed[1])
  text(x = day.count - 6, y = 0.87 * ylim[2], paste(dy, "cases/week\nAvg:", dy %/% 7, "case/day\n", last.progression.pct, "% progress last week"), srt = 20, col = "black")
  
  # Show special measures against propagation.
  ypos = max(country.confirmed, na.rm = TRUE) %/% 3
  if (!is.na(public.spaces.closed.at)) {
    abline(v = public.spaces.closed.at, col = "blue", lw = 1, lty = 5)
    text(x = public.spaces.closed.at - 0.5, y = ypos, "All public spaces are closed", col = "blue", srt = 90)
  }
  if (!is.na(confinement.at)) {
    abline(v = confinement.at, col = "red", lw = 1, lty = 5)
    text(x = confinement.at - 0.5, y = ypos, "Population is confined", col = "red", srt = 90)
  }
  
  legend("topleft", legend = c("Confirmed", "Deaths", "Recovered", "Confirmed trend based on last day", "Confirmed trend based on last week"), col = c("red", "black", "green", "violet", "black"), lty = c(rep(1, 3), 2, 3), lw = c(rep(1, 3), 1.5, 1.5))
  
  diffs <- diff(country.confirmed)
  plot(seq(2, day.count), diffs, type = "h", col = "blue", xlim =  c(day.start + 1, day.count), main = paste("New Confirmed Cases per Day", country.name, sep = " in "), xlab = "Days Since January 21, 2020", ylab = "Number of New Cases per Day", lwd = 5)
  grid()
  
  diffs <- diff(country.deaths)
  plot(seq(2, day.count), diffs, type = "h", col = "black", xlim =  c(day.start + 1, day.count), main = paste("New Deaths per Day", country.name, sep = " in "), xlab = "Days Since January 21, 2020", ylab = "Number of New Deaths per Day", lwd = 5)
  grid()
  
  diffs <- diff(country.recovered)
  plot(seq(2, day.count), diffs, type = "h", col = "green", xlim =  c(day.start + 1, day.count),  main = paste("New Recovered Cases per Day in", country.name, sep = " in "), xlab = "Days Since January 21, 2020", ylab = "Number of New Recovered Cases per Day", lwd = 5)
  grid()
  
  diffs <- diff(country.confirmed - (country.deaths + country.recovered))
  plot(seq(2, day.count), diffs, type = "h", col = "red", xlim =  c(day.start + 1, day.count), main = paste("New Active Cases per Day", country.name, sep = " in "), xlab = "Days Since January 21, 2020", ylab = "Number of New Active Cases per Day", lwd=5)
  grid()
}

covid <- read.csv("csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")
covid.deaths <- read.csv("csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")
covid.recovered <- read.csv("csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")

country.names <- c("Switzerland", "Italy", "France", "Germany", "Austria" , "Spain", "Great Britain", "Denmark", "Norway", "Sweden", "Iceland")
country.count <- length(country.names)
col <- rainbow(country.count)
lty <- seq(1, country.count)
pch <- seq(10, 10 + country.count)

switzerland.confirmed <- get_timeseries_as_vector(subset(covid, Country.Region == "Switzerland"))
switzerland.deaths <- get_timeseries_as_vector(subset(covid.deaths, Country.Region == "Switzerland"))
switzerland.recovered <- get_timeseries_as_vector(subset(covid.recovered, Country.Region == "Switzerland"))

italy.confirmed <- get_timeseries_as_vector(subset(covid, Country.Region == "Italy"))
italy.deaths <- get_timeseries_as_vector(subset(covid.deaths, Country.Region == "Italy"))
italy.recovered <- get_timeseries_as_vector(subset(covid.recovered, Country.Region == "Italy"))

france.confirmed <- get_timeseries_as_vector(subset(covid, Country.Region == "France" & Province.State == ""))
france.deaths <- get_timeseries_as_vector(subset(covid.deaths, Country.Region == "France" & Province.State == ""))
france.recovered <- get_timeseries_as_vector(subset(covid.recovered, Country.Region == "France" & Province.State == ""))

germany.confirmed <- get_timeseries_as_vector(subset(covid, Country.Region == "Germany"))
germany.deaths <- get_timeseries_as_vector(subset(covid.deaths, Country.Region == "Germany"))
germany.recovered <- get_timeseries_as_vector(subset(covid.recovered, Country.Region == "Germany"))

austria.confirmed <- get_timeseries_as_vector(subset(covid, Country.Region == "Austria"))
austria.deaths <- get_timeseries_as_vector(subset(covid.deaths, Country.Region == "Austria"))
austria.recovered <- get_timeseries_as_vector(subset(covid.recovered, Country.Region == "Austria"))

spain.confirmed <- get_timeseries_as_vector(subset(covid, Country.Region == "Spain"))
spain.deaths <- get_timeseries_as_vector(subset(covid.deaths, Country.Region == "Spain"))
spain.recovered <- get_timeseries_as_vector(subset(covid.recovered, Country.Region == "Spain"))

uk.confirmed <- get_timeseries_as_vector(subset(covid, Country.Region == "United Kingdom" & Province.State == ""))
uk.deaths <- get_timeseries_as_vector(subset(covid.deaths, Country.Region == "United Kingdom" & Province.State == ""))
uk.recovered <- get_timeseries_as_vector(subset(covid.recovered, Country.Region == "United Kingdom" & Province.State == ""))

denmark.confirmed <- get_timeseries_as_vector(subset(covid, Country.Region == "Denmark" & Province.State == ""))
denmark.deaths <- get_timeseries_as_vector(subset(covid.deaths, Country.Region == "Denmark" & Province.State == ""))
denmark.recovered <- get_timeseries_as_vector(subset(covid.recovered, Country.Region == "Denmark" & Province.State == ""))

norway.confirmed <- get_timeseries_as_vector(subset(covid, Country.Region == "Iceland"))
norway.deaths <- get_timeseries_as_vector(subset(covid.deaths, Country.Region == "Iceland"))
norway.recovered <- get_timeseries_as_vector(subset(covid.recovered, Country.Region == "Iceland"))

sweden.confirmed <- get_timeseries_as_vector(subset(covid, Country.Region == "Sweden"))
sweden.deaths <- get_timeseries_as_vector(subset(covid.deaths, Country.Region == "Sweden"))
sweden.recovered <- get_timeseries_as_vector(subset(covid.recovered, Country.Region == "Sweden"))

iceland.confirmed <- get_timeseries_as_vector(subset(covid, Country.Region == "Iceland"))
iceland.deaths <- get_timeseries_as_vector(subset(covid.deaths, Country.Region == "Iceland"))
iceland.recovered <- get_timeseries_as_vector(subset(covid.recovered, Country.Region == "Iceland"))

hubei.confirmed <- get_timeseries_as_vector(subset(covid, Country.Region == "China" & Province.State == "Hubei"))
hubei.deaths <- get_timeseries_as_vector(subset(covid.deaths, Country.Region == "China" & Province.State == "Hubei"))
hubei.recovered <- get_timeseries_as_vector(subset(covid.recovered, Country.Region == "China" & Province.State == "Hubei"))

countries.confirmed <- list(switzerland.confirmed, italy.confirmed, france.confirmed, germany.confirmed, austria.confirmed, spain.confirmed, uk.confirmed, denmark.confirmed, norway.confirmed, sweden.confirmed, iceland.confirmed)
countries.deaths <- list(switzerland.deaths, italy.deaths, france.deaths, germany.deaths, austria.deaths, spain.deaths, uk.deaths, denmark.deaths, norway.deaths, sweden.deaths, iceland.deaths)
countries.recovered <- list(switzerland.recovered, italy.recovered, france.recovered, germany.recovered, austria.recovered, spain.recovered, uk.recovered, denmark.recovered, norway.recovered, sweden.confirmed, iceland.recovered)


day.count <- NROW(switzerland.confirmed)
```

# Evolution In Switzerland, Western Europe, and Hubei Province (China)

The figures do not represent the same thing in each country, depending on the extent of the screening tests carried out. Regardless of this problem, the weekly and daily trend curves allow you to get an idea of the foreseeable evolution for the next days (assuming that the screening strategy does not change over time). Respectively, it provides an indication of a possible inflection of the progression of the epidemic in the near future.

## Switzerland
```{r, echo=FALSE, warning=FALSE}
plot_time_series_for_country(switzerland.confirmed, switzerland.deaths, switzerland.recovered, "Switzerland", 55, NA)
```

## Italy
```{r, echo=FALSE, warning=FALSE}
plot_time_series_for_country(italy.confirmed, italy.deaths, italy.recovered, "Italy", NA, 49)
```

## France
```{r, echo=FALSE, warning=FALSE}
plot_time_series_for_country(france.confirmed, germany.deaths, france.recovered, "France", 53, 56)
```

## Germany
```{r, echo=FALSE, warning=FALSE}
plot_time_series_for_country(germany.confirmed, switzerland.deaths, germany.recovered, "Germany", 48, NA)
```

## Austria
```{r, echo=FALSE, warning=FALSE}
plot_time_series_for_country(austria.confirmed, austria.deaths, austria.recovered, "Austria", 52, NA)
```

## Spain
```{r, echo=FALSE, warning=FALSE}
plot_time_series_for_country(spain.confirmed, spain.deaths, spain.recovered, "Spain", NA, 53)
```

## Great Britain
```{r, echo=FALSE, warning=FALSE}
plot_time_series_for_country(uk.confirmed, uk.deaths, uk.recovered, "Great Britain", NA, NA)
```

## Denmark
```{r, echo=FALSE, warning=FALSE}
plot_time_series_for_country(denmark.confirmed, denmark.deaths, denmark.recovered, "Denmark", NA, NA)
```

## Norway
```{r, echo=FALSE, warning=FALSE}
plot_time_series_for_country(norway.confirmed, norway.deaths, norway.recovered, "Norway", NA, NA)
```

## Sweden
```{r, echo=FALSE, warning=FALSE}
plot_time_series_for_country(sweden.confirmed, sweden.deaths, sweden.recovered, "Sweden", NA, NA)
```

## Iceland
```{r, echo=FALSE, warning=FALSE}
plot_time_series_for_country(iceland.confirmed, iceland.deaths, iceland.recovered, "Iceland", NA, NA)
```

## Hubei Province, China
```{r, echo=FALSE, warning=FALSE}
plot_time_series_for_country(hubei.confirmed, hubei.deaths, hubei.recovered, "Hubei Province, China", NA, 2)
```

# Comparison of Number of Confirmed Cases

```{r, echo=FALSE, warning=FALSE}
xrange <- c(35, day.count)
max.y <- 0
for (i in seq(1, country.count)) {
  max.y <- max(c(max.y, countries.confirmed[[i]]), na.rm = TRUE)
}
yrange <- c(0, max.y)
plot(xrange, yrange, main= "Confirmed Cases", type = "n", xlab = "Days Since 2020-01-21", ylab = "Number of Cases")
for (i in seq(1, country.count)) {
  lines(seq(1, day.count), countries.confirmed[[i]], type = "b", col = col[i], lty = lty[i], pch = pch[i], lw = 1)
}
grid()
legend("topleft", legend = country.names, col = col, lty = lty, pch = pch, title = "Country")

max.y <- 0
for (i in seq(1, country.count)) {
  max.y <- max(c(max.y, diff(countries.confirmed[[i]])), na.rm = TRUE)
}
xrange <- c(day.count-10, day.count)
yrange <- c(0, max.y * 1.5)
plot(xrange, yrange, main= "New Confirmed Cases per Day (last 10 days)", type = "n", xlab = "Days Since 2020-01-21", ylab = "Number of New Cases per Day")
for (i in seq(1, country.count)) {
  lines(seq(2, day.count) - 0.45 + 0.08 * i, diff(countries.confirmed[[i]]), type = "h", col = col[i], lty = 1, lw = 4)
}
grid()
legend("topleft", legend = country.names, col = col, lty = 1, lw = 4, title = "Country", ncol = 3)
```

# Comparison of Number of Deaths

```{r, echo=FALSE, warning=FALSE}
xrange <- c(40, day.count)
max.y <- 0
for (i in seq(1, country.count)) {
  max.y <- max(c(max.y, countries.deaths[[i]]), na.rm = TRUE)
}
yrange <- c(0, max.y)
plot(xrange, yrange, main= "Deaths", type = "n", xlab = "Days Since January 21, 2020", ylab = "Number of Deaths")
for (i in seq(1, country.count)) {
  lines(seq(1, day.count), countries.deaths[[i]], type = "b", col = col[i], lty = lty[i], pch = pch[i], lw = 1)
}
grid()
legend("topleft", legend = country.names, col = col, lty = lty, pch = pch, title = "Country")

max.y <- 0
for (i in seq(1, country.count)) {
  max.y <- max(c(max.y, diff(countries.deaths[[i]])), na.rm = TRUE)
}
xrange <- c(day.count-10, day.count)
yrange <- c(0, max.y * 1.5)
plot(xrange, yrange, main= "New Deaths per Day (last 10 days)", type = "n", xlab = "Days Since 2020-01-21", ylab = "Number of New Deaths per Day")
for (i in seq(1, country.count)) {
  lines(seq(2, day.count) - 0.45 + 0.08 * i, diff(countries.deaths[[i]]), type = "h", col = col[i], lty = 1, lw = 4)
}
grid()
legend("topleft", legend = country.names, col = col, lty = 1, lw = 4, title = "Country", ncol = 3)
```

# Comparison of Number of Recovered Cases

WARNING: The figures seem to be wrong on March 23, 2020.

```{r, echo=FALSE, warning=FALSE}
xrange <- c(40, day.count)
max.y <- 0
for (i in seq(1, country.count)) {
  max.y <- max(c(max.y, countries.recovered[[i]]), na.rm = TRUE)
}
xrange <- c(50, day.count)
yrange <- c(0, max.y)
plot(xrange, yrange, main= "Recovered Cases", type = "n", xlab = "Days Since January 21, 2020", ylab = "Number of Cases")
for (i in seq(1, country.count)) {
  d <- nrow(countries.recovered[[i]])
  lines(seq(1, d), countries.recovered[[i]], type = "b", col = col[i], lty = lty[i], pch = pch[i], lw = 1)
}
grid()
legend("topleft", legend = country.names, col = col, lty = lty, pch = pch, title = "Country")

max.y <- 0
for (i in seq(1, country.count)) {
  max.y <- max(c(max.y, diff(countries.recovered[[i]])), na.rm = TRUE)
}
xrange <- c(day.count-10, day.count)
yrange <- c(0, max.y * 1.5)
plot(xrange, yrange, main= "New Recovered Cases per Day (last 10 days)", type = "n", xlab = "Days Since 2020-01-21", ylab = "Number of New Recovered Cases per Day")
for (i in seq(1, country.count)) {
  lines(seq(2, day.count) - 0.45 + 0.08 * i, diff(countries.recovered[[i]]), type = "h", col = col[i], lty = 1, lw = 4)
}
grid()
legend("topleft", legend = country.names, col = col, lty = 1, lw = 4, title = "Country", ncol = 3)
```

# Comparison of Number of Active Cases

WARNING: The figures seem to be wrong on March 23, 2020.

```{r, echo=FALSE, warning=FALSE}
xrange <- c(40, day.count)
max.y <- 0
for (i in seq(1, country.count)) {
  # Rough estmate we take the max of everything and not the max of active case.
  max.y <- max(c(max.y, countries.confirmed[[i]], countries.recovered[[i]], countries.deaths[[i]]), na.rm = TRUE)
}
yrange <- c(0, max.y)
plot(xrange, yrange, main= "Active Cases", type = "n", xlab = "Days Since January 21, 2020", ylab = "Number of Cases")
for (i in seq(1, country.count)) {
  y <- countries.confirmed[[i]] - (countries.recovered[[i]] + countries.deaths[[i]])
  lines(seq(1, d), y, type = "b", col = col[i], lty = lty[i], pch = pch[i], lw = 1)
}
grid()
legend("topleft", legend = country.names, col = col, lty = lty, pch = pch, title = "Country")

min.y <- 0
max.y <- 0
for (i in seq(1, country.count)) {
  y <- countries.confirmed[[i]] - (countries.recovered[[i]] + countries.deaths[[i]])
  max.y <- max(c(max.y, diff(y)), na.rm = TRUE)
  min.y <- min(c(min.y, diff(y)), na.rm = TRUE)
}
xrange <- c(day.count-10, day.count)
yrange <- c(min.y, max.y * 1.5)
plot(xrange, yrange, main= "New Active Cases per Day (last 10 days)", type = "n", xlab = "Days Since 2020-01-21", ylab = "Number of New Active Cases per Day")
for (i in seq(1, country.count)) {
  y <- countries.confirmed[[i]] - (countries.recovered[[i]] + countries.deaths[[i]])
  lines(seq(2, day.count) - 0.45 + 0.08 * i, diff(y), type = "h", col = col[i], lty = 1, lw = 4)
}
grid()
legend("topleft", legend = country.names, col = col, lty = 1, lw = 4, title = "Country", ncol = 3)
```

# About

All the graphs above are generated from the data provided by Johns Hopkins CSSE (provided to the public strictly for educational and academic research purposes).

The source code for generating the graphs is available under MIT License at https://github.com/philippe-docourt/COVID-19/blob/master/covid_19_western_europe.Rmd.

The data are provided as *2019 Novel Coronavirus COVID-19 (2019-nCoV) Data Repository by Johns Hopkins CSSE* (see https://github.com/CSSEGISandData/COVID-19).

See https://github.com/philippe-docourt/COVID-19/blob/master/README.md for more details.

## Data Sources

* World Health Organization (WHO): https://www.who.int/
* DXY.cn. Pneumonia. 2020. http://3g.dxy.cn/newh5/view/pneumonia.
* BNO News: https://bnonews.com/index.php/2020/02/the-latest-coronavirus-cases/
* National Health Commission of the People’s Republic of China (NHC):
* http://www.nhc.gov.cn/xcs/yqtb/list_gzbd.shtml
* China CDC (CCDC): http://weekly.chinacdc.cn/news/TrackingtheEpidemic.htm
* Hong Kong Department of Health: https://www.chp.gov.hk/en/features/102465.html
* Macau Government: https://www.ssm.gov.mo/portal/
* Taiwan CDC: https://sites.google.com/cdc.gov.tw/2019ncov/taiwan?authuser=0
* US CDC: https://www.cdc.gov/coronavirus/2019-ncov/index.html
* Government of Canada: https://www.canada.ca/en/public-health/services/diseases/coronavirus.html
* Australia Government Department of Health: https://www.health.gov.au/news/coronavirus-update-at-a-glance
* European Centre for Disease Prevention and Control (ECDC): https://www.ecdc.europa.eu/en/geographical-distribution-2019-ncov-cases
* Ministry of Health Singapore (MOH): https://www.moh.gov.sg/covid-19
* Italy Ministry of Health: http://www.salute.gov.it/nuovocoronavirus

## Terms of Use

* Data : Copyright 2020 Johns Hopkins University, all rights reserved. See https://github.com/CSSEGISandData/COVID-19/blob/master/README.md.
* Source code for generating graphs: Copyright Philippe Docourt. Licensed under MIT License. See https://github.com/philippe-docourt/COVID-19/blob/master/covid_19_western_europe.Rmd.
* Graphics : Public Domain

